White Wine Quality by Baohua
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(gridExtra)
library(scales)
library(GGally)
library(rpart)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
wwq <- read.csv('wineQualityWhites.csv')
```

**What the data about**:  
The data is from the paper: *Modeling wine preferences by data mining from 
physicochemical properties, by P. Cortez, A. Cerdeira, F. Almeida, 
T. Matos and J. Reis.* It is about the quality of 4898 white wines,
which is rated by at least 3 wine experts with 0 the lowest score and 10 the 
highest score. For each wine, 11 chemical features are collected.

**Variables used in the data and their units**  
Input variables (based on physicochemical tests):  
   1 - fixed acidity (tartaric acid - g / dm^3)  
   2 - volatile acidity (acetic acid - g / dm^3)  
   3 - citric acid (g / dm^3)  
   4 - residual sugar (g / dm^3)  
   5 - chlorides (sodium chloride - g / dm^3  
   6 - free sulfur dioxide (mg / dm^3)  
   7 - total sulfur dioxide (mg / dm^3)  
   8 - density (g / cm^3)  
   9 - pH  
   10 - sulphates (potassium sulphate - g / dm3)  
   11 - alcohol (% by volume)  
Output variable (based on sensory data):   
   12 - quality (score between 0 and 10)  

# Univariate Plots Section

**First several rows of the data**

```{r echo=FALSE, message=FALSE, warning=FALSE, First_rows}
head(wwq)
```

**Dimensionality of the data**

```{r echo=FALSE, message=FALSE, warning=FALSE, Dimension}
dim(wwq)
```

**Structure** 

```{r echo=FALSE, message=FALSE, warning=FALSE, Structure}
str(wwq)
```

**Summary**

```{r echo=FALSE, message=FALSE, warning=FALSE, Summary}
summary(wwq)
```

**Now, let's analyze the data in detail.**

As shown above, there are 4898 items in the data set, and each item is
labeled by the quality level (from 0 to 10), and characterized by 11 features. 
First, I want to plot all the histograms of this 11 features regardless of their
quality level to have a general understanding of the chemical content of white
wines.

```{r echo=FALSE, message=FALSE, warning=FALSE, Hist_all}

# Define a ggplot function that can be applied to all features
ggplot_f1 <- function(feature, xlabel){
  ggplot(aes(x = feature), 
         data = wwq) +
    geom_histogram(bins = 40) +
    xlab(xlabel) +
    theme(text = element_text(size = 8))
}

f1 <- ggplot_f1(wwq$fixed.acidity, 'fixed.acidity (g/dm^3)')
f2 <- ggplot_f1(wwq$volatile.acidity, 'volatile.acidity (g/dm^3)')
f3 <- ggplot_f1(wwq$citric.acid, 'citric.acid (g/dm^3)')
f4 <- ggplot_f1(wwq$residual.sugar, 'residual.sugar (g/dm^3)')
f5 <- ggplot_f1(wwq$chlorides, 'chlorides (g/dm^3)')
f6 <- ggplot_f1(wwq$free.sulfur.dioxide, 'free.sulfur.dioxide (g/dm^3)')
f7 <- ggplot_f1(wwq$total.sulfur.dioxide, 'total.sulfur.dioxide (g/dm^3)')
f8 <- ggplot_f1(wwq$density, 'density (g/cm^3)')
f9 <- ggplot_f1(wwq$pH, 'pH (g/dm^3)')
f10 <- ggplot_f1(wwq$sulphates, 'sulphates (g/dm^3)')
f11 <- ggplot_f1(wwq$alcohol, 'alcohol (% of volumn)')
f12 <- ggplot_f1(wwq$quality, 'quality')

grid.arrange(f12, f1, f2, f3, f4, f5, f6, f7, f8, f9, f10, f11, ncol = 3)
```

There are several interesing histograms. For example, there is huge peak at the
low level of residual.sugar, and the whole histogram is right skewed. To have
a better view, I replot the residual.sugar histogram in log scale:

```{r, echo=FALSE, message=FALSE, warning=FALSE, Hist_residual.sugar_log}
ggplot(aes(x = residual.sugar), 
         data = wwq) +
    geom_histogram(bins = 40) +
    scale_x_log10()
    xlab('residual.sugar (g/dm^3)') +
    theme(text = element_text(size = 8))

```

We can see there is a very obvious bimodal structure, which indicates that the
sugar left after fermentation is likely to be either rather low or rather high,
but very unlikely to be at a moderate level.

For the output variable quality, it is not surprising that only 7 numbers have 
counts, and they are 3, 4, 5, 6, 7, 8 and 9. The mode sits at 6, and wines 
scored 3 and 9 are really rare. If we look at the table for the quality

```{r echo=FALSE, message=FALSE, warning=FALSE, Table_quality}
table(wwq$quality)
```
we can see that there are only 20 wines scored 3, and 5 wines scored 9. I am
very interested in seeing what the differences are between low-quality and
high-quality wines, and for the next plots, I want to do the following 
comparison: wines with score 3 and 4 (low quality) **vs** wines with score 8 
and 9 (high quality), and for each group, there are around 180 samples.

```{r echo=FALSE, message=FALSE, warning=FALSE, FreqPoly_low_high1}

# Create a new dataframe containing only wines with scores 3, 4, 8, and 9
wwq_lh <- wwq[wwq$quality == 3 | wwq$quality == 4 |
                 wwq$quality == 8 | wwq$quality == 9, ]
# Add a new column into the dataframe wwq_lh 'label1'.
wwq_lh$label1 <- ifelse(wwq_lh$quality > 6, 'high quality', 'low quality')  

# Define a ggplot function that can be applied to all features
ggplot_f2 <- function(feature, xlabel){
  ggplot(aes(x = feature, y = ..density.., color = label1), 
         data = wwq_lh) +
    geom_freqpoly(bins = 10) +
    xlab(xlabel) +
    ylab('distribution')
}

f1 <- ggplot_f2(wwq_lh$fixed.acidity, 'fixed.acidity (g/dm^3)')
f2 <- ggplot_f2(wwq_lh$volatile.acidity, 'volatile.acidity (g/dm^3)')
f3 <- ggplot_f2(wwq_lh$citric.acid, 'citric.acid (g/dm^3)')
f4 <- ggplot_f2(wwq_lh$residual.sugar, 'residual.sugar (g/dm^3)')
f5 <- ggplot_f2(wwq_lh$chlorides, 'chlorides (g/dm^3)')
f6 <- ggplot_f2(wwq_lh$free.sulfur.dioxide, 'free.sulfur.dioxide (g/dm^3)')
f7 <- ggplot_f2(wwq_lh$total.sulfur.dioxide, 'total.sulfur.dioxide (g/dm^3)')
f8 <- ggplot_f2(wwq_lh$density, 'density (g/cm^3)')
f9 <- ggplot_f2(wwq_lh$pH, 'pH (g/dm^3)')
f10 <- ggplot_f2(wwq_lh$sulphates, 'sulphates (g/dm^3)')
f11 <- ggplot_f2(wwq_lh$alcohol, 'alcohol (% of volumn)')
f12 <- ggplot_f2(wwq_lh$quality, 'quality')

#extract legend
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

theme_add <- function(fn){
  fn = fn + theme(text = element_text(size = 8),
                  legend.position = 'none')
  return(fn)
}

legend_ex <- g_legend(f1)

grid.arrange(theme_add(f1), theme_add(f2), theme_add(f3), theme_add(f4), 
             theme_add(f5), theme_add(f6), theme_add(f7), theme_add(f8),
             theme_add(f9), theme_add(f10), theme_add(f11), 
             ncol = 3, nrow = 4, 
             legend_ex)
```

It is interesting to note that some of the chemical features are very similar
between high-quality wine and low-quality wine. However, for volatile.acidity, 
citric.acid, free.sulfur.dioxide, pH value, density and alcohol percentage, the
distributions are different between high quality and low quality wines.
The most interesting thing we can notice is that
high-quality wines have higher alcohol percentage than low-quality wines.
Consistently, since alcohol is less dense than water, high-quality wines have a
lower density than low-quality wines. This negative correlation will be further
discussed in bivariate session. 

Now, let's repeat the same high-quality and low-quality analysis, but expand the
sample size to include 5 as low-quality, and 7 as high-quality. In this way,
high quality wines include quality scores 7, 8 and 9; low quality wines include
quality scores 3, 4, and 5.

```{r echo=FALSE, message=FALSE, warning=FALSE, FreqPoly_low_high2}

# Create a new dataframe containing only wines with scores 3, 4, 5, 7, 8, and 9
wwq_lh <- wwq[wwq$quality == 3 | wwq$quality == 4 | wwq$quality == 5 |
              wwq$quality == 7 | wwq$quality == 8 | wwq$quality == 9, ]
# Add a new column into the dataframe wwq_lh 'label1'.
wwq_lh$label1 <- ifelse(wwq_lh$quality > 6, 'high quality', 'low quality')  
wwq_lh$label1 <- factor(wwq_lh$label1)

# Define a ggplot function that can be applied to all features
ggplot_f2 <- function(feature, xlabel){
  ggplot(aes(x = feature, y = ..density.., color = label1), 
         data = wwq_lh) +
    geom_freqpoly(bins = 20) +
    xlab(xlabel) +
    ylab('distribution')
}

f1 <- ggplot_f2(wwq_lh$fixed.acidity, 'fixed.acidity (g/dm^3)')
f2 <- ggplot_f2(wwq_lh$volatile.acidity, 'volatile.acidity (g/dm^3)')
f3 <- ggplot_f2(wwq_lh$citric.acid, 'citric.acid (g/dm^3)')
f4 <- ggplot_f2(wwq_lh$residual.sugar, 'residual.sugar (g/dm^3)')
f5 <- ggplot_f2(wwq_lh$chlorides, 'chlorides (g/dm^3)')
f6 <- ggplot_f2(wwq_lh$free.sulfur.dioxide, 'free.sulfur.dioxide (g/dm^3)')
f7 <- ggplot_f2(wwq_lh$total.sulfur.dioxide, 'total.sulfur.dioxide (g/dm^3)')
f8 <- ggplot_f2(wwq_lh$density, 'density (g/cm^3)')
f9 <- ggplot_f2(wwq_lh$pH, 'pH (g/dm^3)')
f10 <- ggplot_f2(wwq_lh$sulphates, 'sulphates (g/dm^3)')
f11 <- ggplot_f2(wwq_lh$alcohol, 'alcohol (% of volumn)')
f12 <- ggplot_f2(wwq_lh$quality, 'quality')

#extract legend
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

theme_add <- function(fn){
  fn = fn + theme(text = element_text(size = 8),
                  legend.position = 'none')
  return(fn)
}

legend_ex <- g_legend(f1)

grid.arrange(theme_add(f1), theme_add(f2), theme_add(f3), theme_add(f4), 
             theme_add(f5), theme_add(f6), theme_add(f7), theme_add(f8),
             theme_add(f9), theme_add(f10), theme_add(f11), 
             ncol = 3, nrow = 4, 
             legend_ex)
```

This is very similar to the previous analysis. The alcohol
percentage and density are still the most obvious indicators of the wine quality
based on current analysis, especially the skewness of the alcohol distribution,
where high quality wines has a left skewness, and low quality wines has a right
skewness.

Just being curious, I want to add wines with score 6 on top of above figures.

```{r echo=FALSE, message=FALSE, warning=FALSE, FreqPoly_low_mid_high}

# Create a new dataframe
wwq_lh = wwq
# Add a new column into the dataframe wwq_lh 'label1'.
wwq_lh$label1 <- ifelse(wwq_lh$quality > 6, 'high quality', 
                        ifelse(wwq_lh$quality < 6, 'low quality', 'mid quality')
                        )
wwq_lh$label1 <- factor(wwq_lh$label1)

# Define a ggplot function that can be applied to all features
ggplot_f3 <- function(feature, xlabel){
  ggplot(aes(x = feature, y = ..density.., color = label1),
         data = wwq_lh) +
    geom_freqpoly(bins = 20) +
    xlab(xlabel) +
    ylab('distribution')
}

f1 <- ggplot_f3(wwq_lh$fixed.acidity, 'fixed.acidity (g/dm^3)')
f2 <- ggplot_f3(wwq_lh$volatile.acidity, 'volatile.acidity (g/dm^3)')
f3 <- ggplot_f3(wwq_lh$citric.acid, 'citric.acid (g/dm^3)')
f4 <- ggplot_f3(wwq_lh$residual.sugar, 'residual.sugar (g/dm^3)')
f5 <- ggplot_f3(wwq_lh$chlorides, 'chlorides (g/dm^3)')
f6 <- ggplot_f3(wwq_lh$free.sulfur.dioxide, 'free.sulfur.dioxide (g/dm^3)')
f7 <- ggplot_f3(wwq_lh$total.sulfur.dioxide, 'total.sulfur.dioxide (g/dm^3)')
f8 <- ggplot_f3(wwq_lh$density, 'density (g/cm^3)')
f9 <- ggplot_f3(wwq_lh$pH, 'pH (g/dm^3)')
f10 <- ggplot_f3(wwq_lh$sulphates, 'sulphates (g/dm^3)')
f11 <- ggplot_f3(wwq_lh$alcohol, 'alcohol (% of volumn)')
f12 <- ggplot_f3(wwq_lh$quality, 'quality')

#extract legend
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

theme_add <- function(fn){
  fn = fn + theme(text = element_text(size = 8),
                  legend.position = 'none')
  return(fn)
}

legend_ex <- g_legend(f1)

grid.arrange(theme_add(f1), theme_add(f2), theme_add(f3), theme_add(f4), 
             theme_add(f5), theme_add(f6), theme_add(f7), theme_add(f8),
             theme_add(f9), theme_add(f10), theme_add(f11), 
             ncol = 3, nrow = 4,
             legend_ex)
```

The above figure shows very comforting results that the curves for mid
quality wines are in between the curves of high quality and low quality wines,
which are a little bit different from the ones in previous figure. This is 
understandable, since the range on the x axis has changed and we are still
using 20 as the bin number.

To have another view of the data, I scale the x axis to log10 and get the 
follwing figure.

```{r echo=FALSE, message=FALSE, warning=FALSE, FreqPoly_low_mid_high_log}

# Define a ggplot function that can be applied to all features
ggplot_f4 <- function(feature, xlabel){
  ggplot(aes(x = feature, y = ..density.., color = label1),
         data = wwq_lh) +
    geom_freqpoly(bins = 20) +
    scale_x_log10() +
    xlab(xlabel) +
    ylab('distribution')
}

f1 <- ggplot_f4(wwq_lh$fixed.acidity, 'fixed.acidity (g/dm^3)')
f2 <- ggplot_f4(wwq_lh$volatile.acidity, 'volatile.acidity (g/dm^3)')
f3 <- ggplot_f4(wwq_lh$citric.acid, 'citric.acid (g/dm^3)')
f4 <- ggplot_f4(wwq_lh$residual.sugar, 'residual.sugar (g/dm^3)')
f5 <- ggplot_f4(wwq_lh$chlorides, 'chlorides (g/dm^3)')
f6 <- ggplot_f4(wwq_lh$free.sulfur.dioxide, 'free.sulfur.dioxide (g/dm^3)')
f7 <- ggplot_f4(wwq_lh$total.sulfur.dioxide, 'total.sulfur.dioxide (g/dm^3)')
f8 <- ggplot_f4(wwq_lh$density, 'density (g/cm^3)')
f9 <- ggplot_f4(wwq_lh$pH, 'pH (g/dm^3)')
f10 <- ggplot_f4(wwq_lh$sulphates, 'sulphates (g/dm^3)')
f11 <- ggplot_f4(wwq_lh$alcohol, 'alcohol (% of volumn)')
f12 <- ggplot_f4(wwq_lh$quality, 'quality')

#extract legend
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

theme_add <- function(fn){
  fn = fn + theme(text = element_text(size = 8),
                  legend.position = 'none')
  return(fn)
}

legend_ex <- g_legend(f1)

grid.arrange(theme_add(f1), theme_add(f2), theme_add(f3), theme_add(f4), 
             theme_add(f5), theme_add(f6), theme_add(f7), theme_add(f8),
             theme_add(f9), theme_add(f10), theme_add(f11), 
             ncol = 3, nrow = 4,
             legend_ex)
```

One of the figures catch my eye, which is the residual.sugar. For both low
quality and mid quality wines, the densities are bimodal, meaning the sugar left
after fermentation is either low (around 2 g/dm^3) or high (around 10 g/dm^3).
However, for high quality wines, the density of residual.sugar is a little bit
flatter. Another interesing figure is the alcohol one. In the log10 scale, 
high quality is left skewed, low quality is right skewed, and the mid quality
is not skewed.

Specific relations between features and quality level will be further 
investigated in bivariate and multivariate sections, as well as relations 
amoung features.

# Univariate Analysis

### What is the structure of your dataset?

I use white wine quality data. There are 4898 different wines in total, and
their qualities are rated from 0 (worst) to 10 (best). In the data, the quality
scores are 3, 4, 5, 6, 7, 8, 9. With a peak at score 6, the number of wines 
decrease monotonically to lower scores or higher scores.

For each wine, 11 chemical features are measured and reported in numbers with 
certain unit. Most of the features have a pretty standardized histogram, but
the residual.sugar has a very obivious bimodal structure when plotting in log
scale. In addition, the alcohol level has a rather broad distribution.

### What is/are the main feature(s) of interest in your dataset?

The most interesting feature in my dataset is that wines with different
qualities have very different alcohol level distributions. Low quality wines
tend to have a lower alcohol level, and the histogram is right skewed, while 
high quality wines tend to have a higher alcohol level, and the histogram is 
left skewed. Mid quality wines sit in the middle, and the histogram does not
have an obvious skewness.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

The question I am asking here is whether and how the white wine quality is 
predicted by chemical features. Thus, in fact, many features can be used for 
further investigation, since their frequency polygons show a very clear shift
when the wine quality is changing. Just taking total.sulfur.dioxide for example,
when the wine quality is decreasing, the polygons shift from low level to high
level of sulfur dioxide. This is understandable, because higher level of sulfur
dioxide makes the wines taste worse.

### Did you create any new variables from existing variables in the dataset?

Yes, I did. Since for both low quality wines and high quality wines, the sample
size is small, I combined wines with score 3, 4, and 5 into low quality wines, 
and combine wines with scores 7, 8, and 9 into high quality wines. Wines with
score 6 are labeled as mid quality wines. I created a new factor variable called
'label1' to label wines as 'low quality', 'mid quality' and 'high quality' based
on their qualites scores.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Yes, I performed a log transformation on the residual.sugar level. From the
histogram of residual.sugar, I noticed that there is a huge peak for very low
residual.sugar level, and the histogram is skewed. After the log transformation,
a clear bimodal structure appears.

# Bivariate Plots Section

First, I want to have a look at the box plots of wine quality again each of
the features

```{r echo=FALSE, message=FALSE, warning=FALSE, Scatter_quality_vs_all}
# Reorder the factor level of label1
wwq_lh$label1 <- factor(wwq_lh$label1, levels = c('low quality', 
                                                  'mid quality',
                                                  'high quality'))
# Define a ggplot function that can be applied to all features
ggplot_f5 <- function(feature, ylabel){
  ggplot(aes(x = wwq_lh$label1, y = feature), 
         data = wwq_lh) +
    geom_boxplot() +
    scale_y_log10() +
    xlab('quality') +
    ylab(ylabel) +
    theme(text = element_text(size = 8))
}

f1 <- ggplot_f5(wwq$fixed.acidity, 'fixed.acidity \n (g/dm^3)')
f2 <- ggplot_f5(wwq$volatile.acidity, 'volatile.acidity \n (g/dm^3)')
f3 <- ggplot_f5(wwq$citric.acid, 'citric.acid \n (g/dm^3)')
f4 <- ggplot_f5(wwq$residual.sugar, 'residual.sugar \n (g/dm^3)')
f5 <- ggplot_f5(wwq$chlorides, 'chlorides \n (g/dm^3)')
f6 <- ggplot_f5(wwq$free.sulfur.dioxide, 'free.sulfur.dioxide \n (g/dm^3)')
f7 <- ggplot_f5(wwq$total.sulfur.dioxide, 'total.sulfur.dioxide \n (g/dm^3)')
f8 <- ggplot_f5(wwq$density, 'density \n (g/cm^3)')
f9 <- ggplot_f5(wwq$pH, 'pH \n (g/dm^3)')
f10 <- ggplot_f5(wwq$sulphates, 'sulphates \n (g/dm^3)')
f11 <- ggplot_f5(wwq$alcohol, 'alcohol \n (% of volumn)')

grid.arrange(f1, f2, f3, f4, f5, f6, f7, f8, f9, f10, f11, ncol = 3)
```

It can be seen that some of the featues are really good indicators of wine
qualities, and they are residual.sugar, chlorides, total.sulfur.dioxide,
density, pH, and alcohol. In the multivariate section, I will explore more on
how qualities depend on more than one features. Now, let's look at some
relations between different features.

```{r echo=FALSE, message=FALSE, warning=FALSE, density_vs_alcohol}
ggplot(aes(x = alcohol, y = density), 
       data = wwq) +
  geom_point() +
  xlab('alcohol (% of volumn)') +
  ylab('density (g/cm^3)') +
  theme(text = element_text(size = 12))
```

We can see that, besides a strange outlier, there is a clear negative linear 
relation between the density and alcohol level. This result is not surprising, 
since alcohol has a lower density than water, and more alcohol will make the 
density lower.

```{r echo=FALSE, message=FALSE, warning=FALSE, acidity_vs_pH}
f1 <- ggplot(aes(x = pH, y = fixed.acidity), 
             data = wwq) +
        geom_point(alpha = 1/5) +
        xlab('pH') +
        ylab('fixed.acidity (g/dm^3)') +
        theme(text = element_text(size = 12))

f2 <- ggplot(aes(x = pH, y = volatile.acidity), 
             data = wwq) +
        geom_point(alpha = 1/5) +
        xlab('pH') +
        ylab('volatile.acidity (g/dm^3)') +
        theme(text = element_text(size = 12))

f3 <- ggplot(aes(x = pH, y = citric.acid), 
             data = wwq) +
        geom_point(alpha = 1/5) +
        xlab('pH') +
        ylab('citric.acid (g/dm^3)') +
        theme(text = element_text(size = 12))

grid.arrange(f1, f2, f3, ncol = 2)
```

From the figure above, I noticed that fixed.acidity has a clear negative linear
relation with pH level, but this type of relation is not obvious for volatile
acid. For citric acid, there is a decreasing trend, but not very obvious.

```{r echo=FALSE, message=FALSE, warning=FALSE, Sulfur}
ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide), 
       data = wwq) +
    geom_point(alpha = 1/5) +
    xlab('total.sulfur.dioxide \n (g/dm^3)') +
    ylab('free.sulfur.dioxide \n (g/dm^3)') +
    theme(text = element_text(size = 12))
```

We can see that total.sulfur.dioxide and free.sulfur.dioxide has a positive
linear relation.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The feature of interest, which is the wine quality, depends on several other
features. The wine quality increases when pH and alcohol level increases, but
decreases when residual.sugar, chlorides, total.sulfur.dioxide, and density
increases.

Although the median shifts for different quality levels, the general shapes of 
the distributions across different quality levels are very similar, such as the
skewness.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Yes. The density of the wine has a negative linear relation with the alcohol
level, and the total.sulfur.dioxide has a positive linear relation with the
free.sulfur.dioxide.

### What was the strongest relationship you found?

The strongest relationship is between the wine quality and the alcohol level.
Besides this, the density, pH level, residual.sugar, chlorides, and
total.sulfur.dioxide are also correlated strongly with the wine quality. These
features will be used in the predictive model.

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_Alcohol_Density}
ggplot(aes(x = alcohol, y = density, color = label1), 
       data = wwq_lh) +
  geom_point(alpha = 1/5) +
  xlab('alcohol (% of volumn)') +
  ylab('density (g/cm^3)') +
  theme(text = element_text(size = 12))
```

This plot is the same as the one in the last section, but colored based on
their quality level. A clear pattern is shown, where low quality wines
accumulate at the left tail, mid quality wines at middle, and  high quality 
wines at the right tail. 

Let's look at more of this type of figures.

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_pH_fixed.acidity}
ggplot(aes(x = pH, y = fixed.acidity, color = label1), 
       data = wwq_lh) +
  geom_point(alpha = 1/5) +
  xlab('pH') +
  ylab('fixed.acidity (g/dm^3)') +
  theme(text = element_text(size = 12))
```

The pattern in the above figures is not clear, and let's take a log scale

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_pH_fixed.acidity_log}
ggplot(aes(x = pH, y = fixed.acidity, color = label1), 
       data = wwq_lh) +
  geom_point(alpha = 1/5) +
  scale_x_log10() +
  scale_y_log10() +
  xlab('pH') +
  ylab('fixed.acidity (g/dm^3)') +
  theme(text = element_text(size = 12))
```

The pattern is still not very clear, although both pH level and fixed.acidity
show shifts of the median when quality level changes. However, I may still
use this two features in the final classification model.

```{r echo=FALSE, message=FALSE, warning=FALSE, Quality_total.sulfur.dioxide_free.sulfur.dioxide}
ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide, color = label1), 
       data = wwq_lh) +
  geom_point(alpha = 1/5) +
  scale_x_log10() +
  scale_y_log10() +
  xlab('total.sulfur.dioxide (g/dm^3)') +
  ylab('free.sulfur.dioxide (g/dm^3)') +
  theme(text = element_text(size = 12))
```

This shows some vague clusters that separate low quality wines from the other
two. I will try to use this two features in the final model.

```{r Building a Classification Model}
fit <- rpart(I(label1) ~ I(alcohol + density 
                          + log10(total.sulfur.dioxide)
                          + log10(free.sulfur.dioxide)), 
         data = wwq_lh,
         method = 'class')

printcp(fit)
```

From this, we can see that the training error is 0.460 (0.55125 * 0.83556), and 
the cross-validation error is 0.462 (0.55125 * 0.83852). 

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

I tried three pairs of the features. Among them, alcohol and density 
strengthened each other and showed clear clustering of wines based on their
quality. Besides, In log scale, total.sulfur.dioxide and free.sulfur.dioxide 
could also cluster the wines, although not as well as the first pair. However,
the fixed.acidity and pH could not show any structures.


### Were there any interesting or surprising interactions between features?

I really did not expect that the alcohol level has such a strong relation with
wine qualities, and a clear clustering of wines based on their alcohol level and
density is surely surprising to me. However, I am bit confused about why there
is no other strong indicators of the wine quality. Probably, I need to put more
effort in this dataset in order to get more relations out of it.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

Yes, I built a tree model using rpart. Both the traing error and cross
validation error is around 0.46, which is better than pure guess (0.67, I used
3 categories instead of 7). However, this model is pretty preliminary, and I 
don't think I put enough time on this. I need to try more algorithms.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
# Define a ggplot function that can be applied to all features
ggplot_f4 <- function(feature, xlabel){
  ggplot(aes(x = feature, y = ..density.., color = label1),
         data = wwq_lh) +
    geom_freqpoly(bins = 20) +
    scale_x_log10() +
    xlab(xlabel) +
    ylab('distribution')
}

f1 <- ggplot_f4(wwq_lh$fixed.acidity, 'fixed.acidity (g/dm^3)')
f2 <- ggplot_f4(wwq_lh$volatile.acidity, 'volatile.acidity (g/dm^3)')
f3 <- ggplot_f4(wwq_lh$citric.acid, 'citric.acid (g/dm^3)')
f4 <- ggplot_f4(wwq_lh$residual.sugar, 'residual.sugar (g/dm^3)')
f5 <- ggplot_f4(wwq_lh$chlorides, 'chlorides (g/dm^3)')
f6 <- ggplot_f4(wwq_lh$free.sulfur.dioxide, 'free.sulfur.dioxide (g/dm^3)')
f7 <- ggplot_f4(wwq_lh$total.sulfur.dioxide, 'total.sulfur.dioxide (g/dm^3)')
f8 <- ggplot_f4(wwq_lh$density, 'density (g/cm^3)')
f9 <- ggplot_f4(wwq_lh$pH, 'pH (g/dm^3)')
f10 <- ggplot_f4(wwq_lh$sulphates, 'sulphates (g/dm^3)')
f11 <- ggplot_f4(wwq_lh$alcohol, 'alcohol (% of volumn)')
f12 <- ggplot_f4(wwq_lh$quality, 'quality')

#extract legend
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

theme_add <- function(fn){
  fn = fn + theme(text = element_text(size = 8),
                  legend.position = 'none')
  return(fn)
}

legend_ex <- g_legend(f1)

grid.arrange(theme_add(f1), theme_add(f2), theme_add(f3), theme_add(f4), 
             theme_add(f5), theme_add(f6), theme_add(f7), theme_add(f8),
             theme_add(f9), theme_add(f10), theme_add(f11),
             top = 'Distributions of features',
             ncol = 3, nrow = 4,
             legend_ex)
```

### Description One

This figure shows distributions of the 11 features for the three wine qualities,
which is straightforward and informative about the overall relations between
features and wine quality.


### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
ggplot(aes(x = alcohol, y = density), 
       data = wwq) +
  geom_point() +
  xlab('alcohol (% of volumn)') +
  ylab('density (g/cm^3)') +
  theme(text = element_text(size = 12)) +
  ggtitle('density vs alchohol')
```

### Description Two

This is the strongest relation that I can get out of the data, and the two
features alchohol and density are strong indicators of wine qualities.


### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(x = alcohol, y = density, color = label1), 
       data = wwq_lh) +
  geom_point(alpha = 1/5) +
  xlab('alcohol (% of volumn)') +
  ylab('density (g/cm^3)') +
  theme(text = element_text(size = 12)) +
  ggtitle('density vs alchohol with wine quality')
```

### Description Three

This is the same plot as previous one but with colors indicating wine quality.
The three clusters are pretty clearly shown.

------

# Reflection

The dataset contains ratings and 11 chemical featues of 4898 different white 
wines. Myquestion is whether and how the wine quality depends on its chemical 
features. I explored relations between the wine quality and features, as well
as relations among features. At last, I built a predictive tree model to do the
classification.

At the beginning, I saw many features vary their distributions for different
wine qualities. For example, the distribution of the residual.sugar is bimodal
for both mid and low wine qualities, but rather flat for high quality wines. 
However, when I was doing the multivariate analysis, I found that this
difference does not help distinguish wines. What is surprising to me is that 
the alchohol level is a very strong indicator of the wine qualities: good wines
tend to have higher alchohol levels. I am not clear what the reason is, and 
perhaps, wines with more alchohol generally taste better. Besides the alchohol
level, the amount of sulfur dioxide also affects the wine quality. The more
sulfur dioxide, the worse the wine quality is, which makes sense since sulfur
dioxide does not smell good. I built a tree model trying to predict the wine
quality using alchohol level, density, and sulfur dioxide. The model achieves 
an error rate of 0.46, which is better than if one just guesses randomly (in 
this case, the error rate will be 0.67).

Of course, the analysis and modeling so far are still preliminary, and further 
investigation on this dataset is needed. First, I want to know more about the 
background of the data, such as the general experimental procedures and how the 
data were collected. Second, more combinations of the features should be
investigated, which might provide new insights about their relations to the wine
quality. At last, other models should be tested. I used a classification model,
and maybe a regression model could give better results (don't use the 3 
categories I created here, but treat the original ratings as a continuous
variable). 